---
title: "Text Analysis in R"
author: "Tyler George   \nBased on Materials from   \nJulia Silge's Workshop   \n [https://juliasilge.github.io/tidytext-tutorial/site/](https://juliasilge.github.io/tidytext-tutorial/site/)    \n And Data Science in a Box by Mine Çetinkaya-Rundel   \n [https://datasciencebox.org/](https://datasciencebox.org/)"
execute: 
  eval: false
---

# Learning goals

-   Text analysis
-   Sentiment analysis
-   Bing vs Afinn Sentiment Lexicons


## Packages

In addition to `tidyverse` we will be using other packages today

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytext)
library(wordcloud)# Making wordclouds
library(stopwords)
library(googlesheets4)
library(gutenbergr)
gs4_deauth()

```

## Tidytext

-   Using tidy data principles can make many text mining tasks easier, more effective, and consistent with tools already in wide use.
-   Learn more at <https://www.tidytextmining.com/>.

## Follow along

## What is tidy text?

Consider some of the lyrics from **There Is a Light That Never Goes Out** by The Smiths

```{r}
text <- c("Take me out tonight",
          "Where there's music and there's people",
          "And they're young and alive",
          "Driving in your car",
          "I never never want to go home",
          "Because I haven't got one",
          "Anymore")
text
```


This form of the text is not very easy to do analysis with. Think about the data you have been working with this week. 


##################################
QUESTION:
How should data be formatted? 
##################################








#### Tidytext

This will clean (tidy) the data!

```{r}
text_df <- tibble(line = 1:7, text = text)

text_df
```


This is still not quite right. Today, we want a word on each line. This process is called *tokenizing*, where we separate words from their sentences,.


```{r}
text_df |>
  unnest_tokens(word, text)
```












#### What about your favorite songs and song lines? 

I am going to have R read in all of your results from the survey before. 

```{r message=FALSE}
#listening <- read_csv("data/listening.csv")

listening <- 
  read_sheet("https://docs.google.com/spreadsheets/d/1WXKCB_a8AO3CFp49VQ5yTSHYpOfDm1WA9mmYN5jTH1M/edit?usp=sharing") |> 
  janitor::clean_names() |> # re-names the column to be called "songs"
  select(songs) # keeps only the column called songs
  
listening
```


#### Looking for commonalities

```{r}
listening |>
  unnest_tokens(word, songs) |>
  count(word, sort = TRUE)
```



















#### Back to looking for commonalities

```{r}
word_counts<-
  listening |>
  unnest_tokens(word, songs) |>
  anti_join(stop_words) |>                           
  count(word, sort = TRUE)

word_counts
```


#### Visualizing commonalities: bar chart

```{r}
word_counts |>
  ggplot(aes(x = fct_reorder(word, n), y = n)) +
  geom_col() +
  labs(x = "Common words", y = "Count") +
  coord_flip()
```

#### Visualizing commonalities: wordcloud

```{r}
set.seed(1234)
wordcloud(words = word_counts$word, 
          freq = word_counts$n, 
          colors = brewer.pal(5,"Blues"),
          random.order = FALSE)
#Note: You may need to increase the size of your plot area to get this to display properly
```
















## Text Analysis on a Whole Book!

Let's look at a whole book, The Road to Oz by L. Frank Baum.

We will use an open source website of books called Project Gutenberg.


```{r}
full_text <- gutenberg_download(26624) # The number is the ebook number for the book on the website. 
```

Now it's time to *tokenize*, where we separate words from their sentences,  and *tidy* were R puts each word on its own line in a dataset. 

```{r}
tidy_book <- full_text |> # sends the full book 
  mutate(line = row_number()) |> # give s row number to each line
  unnest_tokens(word,text) # separates the words "tokens" onto their own lines
  # This stands for un - nest - tokens

tidy_book
```


#### Find the most common words

```{r}
tidy_book |>
  count(word,sort = T) # this counts and sorts by top counts
```


#### What are the most common words?

Now we want a visualization to compare the common words
```{r}
  tidy_book |>
  count(word) |>
  slice_max(n, n = 20) |>
  ggplot(aes(n, fct_reorder(word, n))) +  
    geom_col()
```


##################################
QUESTION:
Why is this not particularly useful? 

Discuss with your group mates a step we could take before making this graph that would lead to a better, more informative, visualization. 
##################################
ANSWER:



##################################


















#### Stop words

-   In computing, stop words are words which are filtered out before or after processing of text.
-   They usually refer to the most common words in a language, but there is not a single list of stop words used by all natural language processing tools.

#### English stop words

```{r}
get_stopwords()
```

#### Spanish stop words

```{r}
get_stopwords(language = "es")
```


#### Most common non-stopwords 

```{r}
  tidy_book |>
  anti_join(get_stopwords(source = "smart")) |> # This keeps only non-stopwords
  count(word, sort = TRUE) |> #count the frequency of the remaining words
  slice_max(n, n = 20) |> # keep only the top 20 counts
  ggplot(aes(n, fct_reorder(word, n))) +  #start a blank graph
    geom_col() #add a column (bar) graph to the blank graph
```



















#### Sentiment lexicons

One way to think about text is the *sentiment* of words or phrases. These are values that represent positive and negative words and phrases. For example, if someone *died*, that is more negative than say, if someone got a *promotion*. 


Explore some sentiment lexicons, or collections of words and phrases with assigned sentiments. 

```{r}
get_sentiments("bing")
```

```{r}
get_sentiments("afinn")
```



##################################
QUESTION:
How are these two *lexicons* of sentiment different? 
##################################
ANSWER:



##################################









#### Implement sentiment analysis with an `inner_join()` and *bing*

We want a sentiment for each word but only for the words in our book. 

```{r}

tidy_book

tidy_book %>%
  inner_join(get_sentiments("bing")) |> # keeps only words with sentiments in bing
  count(sentiment, sort = TRUE)
```


##################################
QUESTION:
What does your group think the "inner join" does? 
##################################
ANSWER:



##################################













##################################
QUESTION:
Consider the final result. What does this output tell you about the book? 
##################################
ANSWER:



##################################















##################################
QUESTION:
This is very broad information, what else do you want to know about 
   the sentiment of the text? How could we change out approach (not code)
    to get more informative/useful results? 
##################################
ANSWER:



##################################






























#### Analysis with *afinn*

Let's consider the *afinn* lexicon now. 


```{r}
tidy_book |>
  inner_join(get_sentiments("afinn")) |> # get sentiment of each word
  summarize(sum(value))
```


##################################
QUESTION:
When we used *bing* lexicon above we counted the number of positive
  words and the number of negative words. 
  
  This new code adds up the *value* of all the words assigned sentiments. 
  
  For example, the word *happy* has a sentiment of `3`, and *sad* has a 
  sentiment of `-2`, and if we add them, 3 + (-2) we get an overall 
  sentiment of 1. 
  
  
  Compare and contrast the application of *afinn* and *bing*:
  
  a) Which gives more information about your book? 
  
  b) Which is lexicon is more likely to be subjective? Why?
  
  
##################################
ANSWER:





##################################






















## Dive deeper...


#### We can actually get the total contributed afinn sentiment from each word

The top 10 positive words:

```{r}
tidy_book |>
  inner_join(get_sentiments("afinn")) |> # get sentiment of each word
  group_by(word) |>
  summarize(sum_sentiment = sum(value))|> 
  slice_max(sum_sentiment,n=10)
```


This code

1) finds the sentiment for each word (such as *happy* has a sentiment of 3),

2) tells R to sum across each word. If happy was in the text 3 times, it would do 3+3+3 for happy and give it a score of 9.

3) outputs the top 10 words that have contributed the most sentiment in the entire book



Similarly the top 10 most negative words. 

```{r}
tidy_book |>
  inner_join(get_sentiments("afinn")) |> # get sentiment of each word
  group_by(word) |>
  summarize(sum_sentiment = sum(value))|> 
  slice_min(sum_sentiment,n=10)
```












#### Visualize the top positive words


```{r}
set.seed(1234)

top_positives <- 
      tidy_book |>
      inner_join(get_sentiments("afinn")) |> # get sentiment of each word
      group_by(word) |>
      summarize(sum_sentiment = sum(value))|> 
      slice_max(sum_sentiment,n=50) 

wordcloud(words = top_positives$word, 
        freq = top_positives$sum_sentiment, 
        colors = brewer.pal(5,"Blues"),
        random.order = FALSE)
```



#### Visualize the top negative


```{r}

top_negatives <- 
      tidy_book |>
      inner_join(get_sentiments("afinn")) |> # get sentiment of each word
      group_by(word) |>
      summarize(sum_sentiment = sum(value))|> 
      slice_min(sum_sentiment,n=50) 

wordcloud(words = top_negatives$word, 
        freq = abs(top_negatives$sum_sentiment), 
        colors = brewer.pal(5,"Reds"),
        random.order = FALSE)
```




## Your Turn!

What book does *your table* want to analyze today?

1) Go to https://www.gutenberg.org/browse/scores/top

2) Click on a top 100 book or search a book you are interested in. Not all are available and not all will be downloadable. 

3) Replace `26624` below with your own choice. You need the `EBook-No.` from the website. It can be found by scrolling down in the table on the page for the book or in the URL. 



```{r}
full_text <- gutenberg_download(26624) # The number is the ebook number for the book on the website. 
```


#### Tokenize

```{r}
tidy_book <- full_text |> # sends the full book 
  mutate(line = row_number()) |> # give s row number to each line
  unnest_tokens(word,text) # separates the words "tokens" onto their own lines
  # This stands for un - nest - tokens

tidy_book
```


#### Common Words

```{r}
  tidy_book |>
  count(word) |>
  slice_max(n, n = 20) |>
  ggplot(aes(n, fct_reorder(word, n))) +  
    geom_col()
```


#### Most common non-stopwords 

```{r}
  tidy_book |>
  anti_join(get_stopwords(source = "smart")) |> # This keeps only non-stopwords
  count(word, sort = TRUE) |> #count the frequency of the remaining words
  slice_max(n, n = 20) |> # keep only the top 20 counts
  ggplot(aes(n, fct_reorder(word, n))) +  #start a blank graph
    geom_col() #add a column (bar) graph to the blank graph
```

#### Sentiment Analysis with Bing

```{r}
tidy_book %>%
  inner_join(get_sentiments("bing")) |> # keeps only words with sentiments in bing
  count(sentiment, sort = TRUE)
```


#### Sentiment Analysis with Afinn

```{r}
tidy_book |>
  inner_join(get_sentiments("afinn")) |> # get sentiment of each word
  summarize(sum(value))
```

#### Top 10 words that contributed the most positive sentiment
```{r}
tidy_book |>
  inner_join(get_sentiments("afinn")) |> # get sentiment of each word
  group_by(word) |>
  summarize(sum_sentiment = sum(value))|> 
  slice_max(sum_sentiment,n=10)
```

#### Top 10 words that contributed the most negative sentiment

```{r}
tidy_book |>
  inner_join(get_sentiments("afinn")) |> # get sentiment of each word
  group_by(word) |>
  summarize(sum_sentiment = sum(value))|> 
  slice_min(sum_sentiment,n=10)
```




#### Visualize the top positive words


```{r}
set.seed(1234)

top_positives <- 
      tidy_book |>
      inner_join(get_sentiments("afinn")) |> # get sentiment of each word
      group_by(word) |>
      summarize(sum_sentiment = sum(value))|> 
      slice_max(sum_sentiment,n=50) 

wordcloud(words = top_positives$word, 
        freq = top_positives$sum_sentiment, 
        colors = brewer.pal(5,"Blues"),
        random.order = FALSE)
```



#### Visualize the top negative


```{r}

top_negatives <- 
      tidy_book |>
      inner_join(get_sentiments("afinn")) |> # get sentiment of each word
      group_by(word) |>
      summarize(sum_sentiment = sum(value))|> 
      slice_min(sum_sentiment,n=50) 

wordcloud(words = top_negatives$word, 
        freq = abs(top_negatives$sum_sentiment), 
        colors = brewer.pal(5,"Reds"),
        random.order = FALSE)
```

